Mixed effect tweedie model in TMB

### 2019-04-18 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
add 'day of event' SST to temperature models; not appearing very useful... what about trying a mean from the month prior to the survey?

```{r setup}
library(tidyverse)
library(sdmTMB)
library(future)
plan(multiprocess)
d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
#d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
d_trawl$depth_scaled3 <- (d_trawl$depth_scaled)^3
unique(d_trawl$year)
```

### Functions for model exploration:
```{r}
aic <- function(m) {
  k <- length(m$model$par)
  nll <- m$model$objective
  2 * k - 2 * (-nll)
}

generate_cv_args <- function(
                formula = formula,
                time_varying = time_varying,
                ar1_fields = c(FALSE), # simiple, fast defaults
                include_spatial = c(FALSE), # simiple, fast defaults
                n_knots = c(100), # simiple, fast defaults
                seed = c(999)
  ){
  arguments <- expand.grid(
    formula = formula,
    time_varying = time_varying,
    include_spatial = include_spatial,
    ar1_fields = ar1_fields,
    n_knots = n_knots,
    seed = seed
  )
  #browser()
  as.data.frame(arguments)
}

sdmTMB_cv_batch <- function (cv_args, data, ...) {
  cv_batch <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl, ...)
  cv_out <- cv_args
  cv_out$cv_loglik <- c()
  cv_out$all_converged <- c()
  cv_out$max_grad <- c()
  for (i in seq_len(nrow(cv_args))) {
    cv_out$cv_loglik[[i]] <- round(cv_batch[[i]]$sum_loglik, 1)
    cv_out$all_converged[[i]] <- all(cv_batch[[i]]$converg)
    cv_out$max_grad[[i]] <- max(cv_batch[[i]]$max_grad)
    if (is.null(cv_out$time_varying[[i]])) 
      cv_out$time_varying[[i]] <- as.character(NA)
  }
  cv_out$formula <- as.character(cv_out$formula)
  cv_out$time_varying <- as.character(cv_out$time_varying)
  list(summary = as.data.frame(cv_out), cv_model_set = cv_batch)
}

plot_cv_meshes <- function(sdmTMB_cv_output) {
  cv_output <- sdmTMB_cv_output
  k <- max(cv_output[["data"]]$cv_fold)
  if (k == 2 | k == 5 | k == 10) {
    op <- graphics::par(
      mfrow = c((sqrt(k)), ceiling(sqrt(k))),
      mar = c(1, 1, 1, 1)
      )
  } else { 
    op <- graphics::par(
      mfrow = c(ceiling(sqrt(k)), ceiling(sqrt(k))),
      mar = c(1, 1, 1, 1)
    )
  }
  for (i in seq_len(k)){
    plot_spde(cv_output$models[[i]]$spde)
  }
  graphics::par(op)
}

```

Short example: 
```{r, eval=FALSE}
cv_args <- generate_cv_args (formula = list(
    temperature_c ~ 0 + as.factor(year) + depth_scaled),
    time_varying = list(NULL),
    ar1_fields = c(FALSE), 
    include_spatial = c(FALSE),
    n_knots = c(30), 
    seed = c(111,222)
  )

cv <- sdmTMB_cv_batch(cv_args, d_trawl, k_folds = 4, time = "year", x = "X", y = "Y")
glimpse(cv$summary)
plot_cv_meshes(cv$cv_model_set[[1]])
plot_cv_meshes(cv$cv_model_set[[2]])
```

@@

## Just compare fixed slope models with simplest random effect structure
```{r}
cv_args <- generate_cv_args (formula = list(
    temperature_c ~ 0 + as.factor(year) + depth_scaled,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3),
    time_varying = list(NULL),
    ar1_fields = c(FALSE), 
    include_spatial = c(FALSE),
    n_knots = c(150), #start with close to projected optimum n_knots 
    seed = c(1111,2221,3331)
  )

@@

## Compare time-varying slope models still with simplest random effect structure
```{r}
cv_args <- generate_cv_args (formula = list(
    temperature_c ~ 0 + as.factor(year)),
    time_varying = list(
    ~ 0 + depth_scaled,
    ~ 0 + depth_scaled + depth_scaled2,
    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
    ar1_fields = c(FALSE), 
    include_spatial = c(FALSE),
    n_knots = c(150), # keep same n_knots
    seed = c(1111,2221,3331) # keep same seeds
    )

@@

## Now vary n_knots and random effect configuation for top model(s) 
```{r}
cv_args <- generate_cv_args(formula = list(
    temperature_c ~ 0 + as.factor(year)),
    time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2,
    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
    ar1_fields = c(FALSE, TRUE), 
    include_spatial = c(FALSE, TRUE),
    n_knots = c(85,110,135),
    seed = c(1111,2221,3331)
    )

@@

## Change seeds 
```{r}
cv_args <- generate_cv_args(formula = list(
    temperature_c ~ 0 + as.factor(year)),
    time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2,
    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
    ar1_fields = c(FALSE), 
    include_spatial = c(TRUE),
    n_knots = c(120),
    seed = c(4441,5551,6661,7771,8881,9991)
    )

@@

Load saved batches:
```{r}
# non_time_varying <- readRDS(file = "cv-non-time-varying.rds")
# time_varying <- readRDS(file = "cv-time-varying.rds")
# vary_rf <- readRDS(file = "cv-varying-rf.rds")
# knot_refining <- readRDS(file = "cv-refine-knots.rds")

# non_time_varying <- read.csv(file = "cv-non-time-varying.csv")
# time_varying <- read.csv(file = "cv-time-varying.csv")
# vary_rf <- read.csv(file = "cv-varying-rf.csv")
# knot_refining <- read.csv(file = "cv-refine-knots.csv")

all_cv_loglik <- rbind(non_time_varying$summary, time_varying$summary, non_tv$summary, tvar$summary, vary_rf$summary, vary_rf_knots$summary, knot_refining$summary, more_seeds$summary, deparse.level = 1, make.row.names=TRUE)
View(all_cv_loglik)

@@

Current top model
```{r}
plan(multiprocess)
temp3_rw <- sdmTMB_cv(data = d_trawl, n_knots = 130, x = "X", y ="Y", k_folds = 10, 
                   formula = temperature_c ~ 0 + as.factor(year),
                   time_varying = ~ 0 + depth_scaled + depth_scaled2,
                   time = "year", family = gaussian(link = "identity"), 
                   ar1_fields = FALS, include_spatial = TRUE)
temp3_rw$models
sum(temp3_rw$sum_loglik)
```

@@

```{r}
m_temp3_rw <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year),
  time_varying =  ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
  ar1_fields = TRUE, include_spatial = TRUE, 
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions3_rw <- predict(m_temp3_rw)
predictions3_rw$data$residuals <- residuals(m_temp3_rw)
cor(predictions3_rw$data$est, predictions3_rw$data$temperature_c)^2 # 
aic(m_temp3_rw)
```

@@

## Plotting predictions
```{r}
qcs_grid$depth_scaled3 <- (qcs_grid$depth_scaled)^3
predictions_nd <- predict(m_temp3_rw, newdata = qcs_grid)
#saveRDS(predictions3_rw_nd$data, file = "qcs-temp.rds")
```

@@

```{r, eval=FALSE}
plot_map_raster(predictions_nd$data, "est_re_s") +
  ggtitle("Spatial random effects only") +
  scale_fill_gradient2()
```

```{r}
plot_map_raster(predictions_nd$data, "est_re_st") +
  ggtitle("Spatiotemporal random effects only") +
  scale_fill_gradient2()
```

```{r}
ggplot(predictions3_rw$data, aes(temperature_c, est)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  xlim(3,15) +
  ylim(3,15) +
  geom_abline()
```

>>>>

```{r setup}
library(tidyverse)
library(sdmTMB)
library(future)
library(dplyr)

plan(multiprocess)
# d_trawl1 <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
# nrow(d_trawl1)
d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/dat-sensor-trawl-SST.rds"))
glimpse(d_trawl)
nrow(d_trawl)
head(d_trawl)
# d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
d_trawl <- mutate(d_trawl,
  depth_scaled = arm::rescale(log(depth_m)),
  depth_scaled2 = depth_scaled^2,
  depth_scaled3 = depth_scaled^3
)
unique(d_trawl$year)
d_trawl[is.na(d_trawl$year), ]$year <- 2003
```

### Functions for model exploration:
```{r}
# aic <- function(m) {
#   k <- length(m$model$par)
#   nll <- m$model$objective
#   2 * k - 2 * (-nll)
# }

generate_cv_args <- function(
                             formula = formula,
                             time_varying = time_varying,
                             ar1_fields = c(FALSE), # simiple, fast defaults
                             include_spatial = c(FALSE), # simiple, fast defaults
                             n_knots = c(100), # simiple, fast defaults
                             seed = c(999)) {
  arguments <- expand.grid(
    formula = formula,
    time_varying = time_varying,
    include_spatial = include_spatial,
    ar1_fields = ar1_fields,
    n_knots = n_knots,
    seed = seed
  )
  as.data.frame(arguments)
}

sdmTMB_cv_batch <- function(cv_args, data, ...) {
  cv_batch <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl, ...)
  cv_out <- cv_args
  cv_out$cv_loglik <- c()
  cv_out$all_converged <- c()
  cv_out$max_grad <- c()
  for (i in seq_len(nrow(cv_args))) {
    cv_out$cv_loglik[[i]] <- round(cv_batch[[i]]$sum_loglik, 1)
    cv_out$all_converged[[i]] <- all(cv_batch[[i]]$converged)
    cv_out$max_grad[[i]] <- max(cv_batch[[i]]$max_grad)
    if (is.null(cv_out$time_varying[[i]])) {
      cv_out$time_varying[[i]] <- as.character(NA)
    }
  }
  cv_out$formula <- as.character(cv_out$formula)
  cv_out$time_varying <- as.character(cv_out$time_varying)

  cv_out <- cv_out %>%
    group_by(seed) %>%
    mutate(., max_ll_per_seed = max(cv_loglik)) %>%
    mutate(., delta_ll = cv_loglik - max_ll_per_seed) %>%
    ungroup()
  row.names(cv_out) <- NULL
  list(summary = as.data.frame(cv_out), cv_model_set = cv_batch)
}

# # this code keeps breaking... working, fails, worked again, then failed again...
# plot_cv_meshes <- function(sdmTMB_cv_output) {
#   cv_output <- sdmTMB_cv_output
#   k <- max(cv_output$data$cv_fold) # should probably rename cv_fold to fold_id
#   if (k == 2 | k == 5 | k == 10) {
#     op <- graphics::par(
#       mfrow = c((sqrt(k)), ceiling(sqrt(k))),
#       mar = c(1, 1, 1, 1)
#       )
#   } else {
#     op <- graphics::par(
#       mfrow = c(ceiling(sqrt(k)), ceiling(sqrt(k))),
#       mar = c(1, 1, 1, 1)
#     )
#   }
#   for (i in seq_len(k)){
#     plot_spde(cv_output$models[[i]]$spde)
# #FIXME: Error in xy.coords(x, y, xlabel, ylabel, log) :
# #  'x' is a list, but does not have components 'x' and 'y'
#   }
#   graphics::par(op)
# }
```

Short example: 
```{r, eval=FALSE}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled + depth_scaled2 + depth_scaled3
  ),
  time_varying = list(NULL),
  ar1_fields = c(FALSE),
  include_spatial = c(FALSE),
  n_knots = c(135, 140, 145, 150),
  seed = c(1)
)

cv <- sdmTMB_cv_batch(cv_args, d_trawl, k_folds = 4, time = "year", x = "X", y = "Y")
glimpse(cv$summary)
write.csv(cv$summary, file = "cv-test.csv")
cv_test <- read_csv(file = "cv-test.csv")
glimpse(cv)
View(cv$summary)
# plot_cv_meshes(cv$cv_model_set[[1]])
# plot_cv_meshes(cv$cv_model_set[[2]])
```

@@

## Just compare fixed slope models with simplest random effect structure
```{r}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 1 + SST + depth_scaled,
    temperature_c ~ 1 + SST + depth_scaled + depth_scaled2,
    temperature_c ~ 1 + SST + depth_scaled + depth_scaled2 + depth_scaled3,
    temperature_c ~ 0 + as.factor(year) + depth_scaled,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled,
    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled + depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled + depth_scaled2 + depth_scaled3,
    temperature_c ~ 0 + as.factor(year) + SST * depth_scaled,
    temperature_c ~ 0 + as.factor(year) + SST * depth_scaled + SST * depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + SST * depth_scaled + SST * depth_scaled2 + SST * depth_scaled3
  ),
  time_varying = list(NULL),
  ar1_fields = c(FALSE),
  include_spatial = c(FALSE),
  n_knots = c(100),
  seed = c(1111, 2222, 3333, 4444)
)

@@

## Compare time-varying slope models still with simplest random effect structure
All of top models for any seed contained year and non showed a linear effect of depth
```{r}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + as.factor(year) + SST,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2,
    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3
  ),
  ar1_fields = c(FALSE),
  include_spatial = c(FALSE),
  n_knots = c(100), # keep same n_knots
  seed = c(1111, 2222, 3333, 4444) # keep same seeds
)

@@

## Now vary random effect configuation for top model(s) 
```{r}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + as.factor(year) + SST,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE, TRUE),
  include_spatial = c(FALSE, TRUE),
  n_knots = c(100),
  seed = c(1111, 2222, 3333, 4444)
)

@@

## Change seeds 
```{r}
cv_args <- generate_cv_args(
    formula = list(
    temperature_c ~ 0 + as.factor(year) + SST,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE, TRUE),
  include_spatial = c(TRUE),
  n_knots = c(120),
  seed = c(5555, 6666, 7777, 8888, 9999)
)

@@

Load saved batches:
```{r, eval=FALSE}
# non_time_varying <- readRDS(file = "cv-non-time-varying.rds")
# time_varying <- readRDS(file = "cv-time-varying.rds")
# vary_rf <- readRDS(file = "cv-varying-rf.rds")
# knot_refining <- readRDS(file = "cv-refine-knots.rds")

non_time_varying <- read.csv(file = "cv-non-time-varying.csv")
time_varying <- read.csv(file = "cv-time-varying.csv")
vary_rf <- read.csv(file = "cv-varying-rf.csv")
non_time_varying1 <- read.csv(file = "cv-non-time-varying1.csv")
time_varying1 <- read.csv(file = "cv-time-varying1.csv")
vary_rf1 <- read.csv(file = "cv-varying-rf1.csv")
refine_knots1 <- read.csv(file = "cv-refine-knots1.csv")
more_seeds <- read.csv(file = "cv-more-seeds.csv")
more_seeds1 <- read.csv(file = "cv-more-seeds1.csv")

all_cv_loglik <- rbind(
  non_time_varying,
  time_varying,
  vary_rf,
  non_time_varying1,
  time_varying1,
  vary_rf1,
  refine_knots1,
  more_seeds,
  more_seeds1,
  deparse.level = 1, make.row.names = TRUE
)

@@

Current top model
```{r}
plan(multiprocess)
temp3_rw <- sdmTMB_cv(
  data = d_trawl, n_knots = 130, x = "X", y = "Y", k_folds = 10,
  formula = temperature_c ~ 0 + as.factor(year) + SST,
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", family = gaussian(link = "identity"),
  ar1_fields = FALSE, include_spatial = TRUE, seed = 999
)
# temp3_rw$models
sum(temp3_rw$sum_loglik)
```

@@

```{r}
m_temp3_rw <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year) + SST,
  time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

predictions3_rw <- predict(m_temp3_rw)
predictions3_rw$data$residuals <- residuals(m_temp3_rw)
cor(predictions3_rw$data$est, predictions3_rw$data$temperature_c)^2 #
# aic(m_temp3_rw)
```

@@

## Plotting predictions
#FIXME: DON'T HAVE SST TEMP FOR PREDICITON GRID YET...

```{r}
qcs_grid$depth_scaled3 <- (qcs_grid$depth_scaled)^3
predictions_nd <- predict(m_temp3_rw, newdata = qcs_grid)
# saveRDS(predictions3_rw_nd$data, file = "qcs-temp.rds")
```

@@

```{r, eval=FALSE}
plot_map_raster(predictions_nd$data, "est_re_s") +
  ggtitle("Spatial random effects only") +
  scale_fill_gradient2(low = "navy", mid = "white", high = "darkred")
```

```{r}
plot_map_raster(predictions_nd$data, "est_re_st") +
  ggtitle("Spatiotemporal random effects only") +
  scale_fill_gradient2(low = "navy", mid = "white", high = "darkred")
```

```{r}
ggplot(predictions3_rw$data, aes(temperature_c, est)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  xlim(3, 15) +
  ylim(3, 15) +
  geom_abline()
```
